---
title: "Take-home Exercise 1 - Data Visualizations in Singapore Residence by Planning Area"
author: "PENG Hsiao Yuan"
---
# 1. Overview

# 1.1 Background Information and Data Source

The data set is downloaded from "https://www.singstat.gov.sg/"

File Name: "Singapore Residents by Planning Area / Subzone, Single Year of Age and Sex, June 2024"

# 1.2 Importing Data and Install Packages 

Loading and Installing packages

```{r}
pacman::p_load(tidyverse, haven, knitr,
               patchwork, ggthemes, scales,
               ggridges, ggpubr, gganimate,
               ggdist, ggtext, ggalt,
               cowplot, ggnewscale)
```

```{r}
```

Importing Data

```{r}
exam_data <- read.csv("respopagesexfa2024/respopagesexfa2024.csv")
```

# 2. Background and Notes 

A local online media company that publishes daily content on digital platforms is planning to release an article on demographic structures and distribution of Singapore in 2024.

# 2.1 Task Objective

Assuming the role of the graphical editor of the media company, you are tasked to prepare at most three data visualisation for the article.

# 2.2 Column Headers:

```{r}
df <- tibble(
  PA = c("Planning Area"),
  SZ = c("Subzone"),
  Sex = c("Sex"),
  FA = c("Age Group"),
  Pop = c("Floor Area of Residence"),
  Time = c("Resident Count(Population)")
)

df %>%
  knitr::kable(caption = "Column Name") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    full_width = FALSE
  )
```

Note:

1)
For June 2024, Planning Areas refer to areas demarcated in the Urban Redevelopment Authority's Master Plan 2019.

2) 
Data from 2003 onwards exclude residents who have been away from Singapore for a continuous period of 12 months or longer as at the reference period.

3) 
The figures have been rounded to the nearest 10. 

4) 
The data may not add up due to rounding.











